Improved state-space models for inference about spatial and temporal variation in abundance from count data

### 2012-10-02 Dail_Mad_ext.tex
discussion

Our emphasis was on increasing the practical utility of this class of
models, and so we avoided several conceptually-interesting extensions
that we belive would be computationally prohibitive in many
cases. Nonetheless, we will discuss one---spatially-explict models of
immigration.

In our models immigration is currently modeled as independent of
population sizes in other sites. Several alternatives suggest
themselves if we think about movement in a metapopulation
context ([cite]): immigration depending on total or mean
abundance across sites the previous time step; immigration
depending on abundance in nearby sites only; and immigration
depending on abundance at sites as a function of their distance
from the receiving site (LOTS OF CITATIONS). These models would
be based on the assumptions that either the study sites
spatially cover the whole range of the metapopulation or the
annual changes in the populations sampled are representative of
a larger metapopulation. Initial tests suggest that it would be
possible to fit some or all of these models, at least in the
Bayesian framework.

@@

The modeling framework we described can be used to address many of the
most pressing issue in ecology and conservation biology. For example,
it is possible to test hypotheses about temporal and spatial
population regulation. Furthermore, the Bayesian approach is very
useful in that it can be used to combine multiple sources of data to
develop mechanistic models of population dynamics. Thus one can test
hypotheses about the effects of climate change on either explicit
demographic parameters or in derived parameters such as population
growth rate. Furthermore, under the Bayesian approach, population
viability analysis is trivial because projecting populations into the
future can be done as a component of the MCMC analysis. This allows
for the computation of posterior distributions of parameters such as
quasi-extinction probability.

@@

We thank J. Andrew Royle and T. Scott Sillett for helpful comments
throughout this process.

>>>>

Our emphasis was on increasing the practical utility of this class of
models, and so we avoided several conceptually-interesting extensions
that we belive would be computationally prohibitive in many
cases. Nonetheless, we will discuss one---spatially-explict models of
immigration.
In our models immigration is currently modeled as independent of
population sizes in other sites. Several alternatives suggest
themselves if we think about movement in a metapopulation
context ([cite]): immigration depending on total or mean
abundance across sites the previous time step; immigration
depending on abundance in nearby sites only; and immigration
depending on abundance at sites as a function of their distance
from the receiving site (LOTS OF CITATIONS). These models would
be based on the assumptions that either the study sites
spatially cover the whole range of the metapopulation or the
annual changes in the populations sampled are representative of
a larger metapopulation. Initial tests suggest that it would be
possible to fit some or all of these models, at least in the
Bayesian framework.

@@

The modeling framework we described can be used to address many
pressing issues in ecology and conservation biology. For example,
%it is possible to test hypotheses about temporal and spatial
%population regulation. Furthermore, the Bayesian approach is very
%useful in that it can be used to combine multiple sources of data to
%develop mechanistic models of population dynamics. Thus one can test
it allows for inference %hypotheses
about the effects of climate change on either explicit
demographic parameters or in derived parameters such as population
growth rate. Furthermore, %under the Bayesian approach, population
in addition to assessing past climate change on population parameters,
projecting populations under future climate scenarios is also
possible. Thus, we expect this modeling methodology can be used to
make quantiative assessments of species vulnerability to climate
change and other important human-induced drivers of population dynamics.

@@

We thank J. Andrew Royle and T. Scott Sillett for helpful comments
and suggestions.
